---
title: "Hw2"
author: "Hubert Chang"
date: "2026-10-06"
output: github_document
---

#Problem 0 set up
```{r}
library (tidyverse)
library (readxl)
```

#Problem 1
```{r}
nytransit = read.csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names()|>
  select (line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada) |>
  
  mutate(
    entry = recode_values(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )

```
The dataset contains information on the subway line, station name, the station's latitude and longitude, the routes from 1 to 11, the type of entrances, and if you can have entry, whether it has a vending machine and if it is ada compliant.
The data have 1868 observations with 19 variables. 
The data can be tidied up more because route 1 - 11 all have their own column with a lot of missing data. 

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.

```{r}
nytransit |>
  distinct(line, station_name) |>
  nrow()
```
There are 465 distinct stations.

How many stations are ADA compliant?
```{r}
nytransit |>
  distinct(line, station_name, ada) |>
  filter (ada == TRUE) |>
  nrow()
```
84 stations are ADA compliant


What proportion of station entrances / exits without vending allow entrance?
```{r}
nytransit|>
  filter( vending == "NO" ) |>
  pull (entry) |>
  mean()
```
About 37.70 % of the station entrance/ exits is without a vending


#Reforming data 

```{r}
nytransit_clean =
  nytransit |>
  mutate (
    across (
      route1:route11, as.character)) |>
  pivot_longer(
    cols = route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )
```

How many distinct stations serve the A train? 
```{r}
nytransit_clean |>
  filter(route_name == "A") |>
  distinct(line, station_name)|>
  nrow()
```
60 distinct stations

Of the stations that serve the A train, how many are ADA compliant?

```{r}
nytransit_clean |>
  filter(route_name == "A", ada == "TRUE") |>
  distinct(line, station_name)|>
  nrow()
```
17 stations are ADA compliant.

#Problem 2

Mr Trash Wheel
```{r}
trashwheel =
  read_excel("data/202610 Trash Wheel Collection Data.xlsx",
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N"),
            skip = 1,
            na = c("NA", ".", "") 
            ) |>
  janitor :: clean_names() |>
  drop_na(dumpster) |>
  mutate(
    sports_balls = as.integer (
      round(sports_balls)
    )) |>
  
  mutate(
    year = as.numeric(year), wheel_name = "Mr. Trash Wheel"
  )
```
Professor Trash Wheel
```{r}
profwheel = read_excel("data/202610 Trash Wheel Collection Data.xlsx",
            sheet = "Professor Trash Wheel",
            skip = 1,
            na = c("NA", ".", "") 
            ) |>
  janitor :: clean_names() |>
  drop_na(dumpster) |>

 mutate(
    year = as.numeric(year), wheel_name = "Professor Trash Wheel" 
  ) |>
  
  mutate(
    plastic_bottles = as.numeric(plastic_bottles)
  )
```

Gwynnda Trash Wheel

```{r}
gwynndawheel = read_excel("data/202610 Trash Wheel Collection Data.xlsx",
            sheet = "Gwynnda the Good Wheel of the W",
            skip = 1,
            na = c("NA", ".", "") 
            ) |>
  janitor :: clean_names() |>
  drop_na(dumpster) |>
  
  mutate(
    year = as.numeric(year), wheel_name = "Gwynnda"
  )
```

Combining the three
```{r}
trashwheel_tidy =
  bind_rows (trashwheel, profwheel, gwynndawheel) |>
  janitor::clean_names() |>
  relocate (wheel_name)
```


The combined trash wheel have `r nrow(trashwheel_tidy)` observations. It also have `r ncol(trashwheel_tidy)` variables. Main variables in the document contains wheel_name, dumpster, month, year, date, weight_tons, volume_cubic_yards, plastic_bottles,polystyrene, cigarette_butts, glass_bottles, plastic_bags, wrappers, sports_balls, homes_powered

```{r}
prof_weight = 
  trashwheel_tidy |>
  filter (wheel_name == "Professor Trash Wheel") |>
  pull (weight_tons) |>
  sum (na.rm = TRUE)
```


Professor Trash Wheel collected `r round(prof_weight)` tons
```{r}
gwynnda_cig = 
  trashwheel_tidy |>
  filter(wheel_name == "Gwynnda", month == "June", year == 2022) |>
  pull(cigarette_butts) |>
  sum (na.rm = TRUE)
```
  

Gwynnda collected `r format(gwynnda_cig) ` cigarette butts in June of 2022


#Problem 3

```{r}
mci_df= read.csv("data_mci/data_mci/MCI_baseline.csv",
  na = c("NA", ".", ""),
  skip = 1 
                ) |>
janitor :: clean_names() 
```

```{r}
mci_data = 
  mci_df |>
  
mutate(
  sex = recode_values(
      sex,
      1 ~ "Male",
      0 ~ "Female"
    ),
  apoe4 = recode_values(
      apoe4,
      1 ~ "APOE4 Carrier",
      0 ~ "APOE4 NON-Carrier"
    )
      )  |> #converting from numbers to characters

filter(is.na(age_at_onset | age_at_onset > current_age)) #removing MCI onset before study baseline
```

The baseline file had a note above, so we have to remove that with skip = 1, I also removed any data that had response with NA, . , or blank. And then I recoded the values from numbers in sex and apoe4 from 1/0 to characters (male/female) (apoe carrier or not).Lastly, I removed anyone who had MCI onset before basline age.

```{r}
mci_women = 
  mci_data |>
  filter (sex == "Female") |>
  mutate ( carrier = apoe4 == "APOE4 Carrier") |>
  pull (carrier) |>
  mean()
```

`r nrow(mci_df)` participants recruited and of those `r filter(mci_df, !is.na(age_at_onset)) |> nrow()` developed MCI. The average baseline age is `r mean(pull(mci_data, current_age))` years old. Among the study of women about `r round(100*mci_women)`% are APOE4 carriers


```{r}
amyloid_df = read.csv("data_mci/data_mci/mci_amyloid.csv",
  na = c("NA", ".", ""),
  skip = 1) |>
  janitor:: clean_names() |>
  rename(id = study_id) 
```

I pulled the data from the mci_amyloid.csv file, then I removed any missing data that had "NA", ".", or blank. I also had to skip the first line because it is just headder. Then I used janitor to clean the file and then renamed study_id to id so it matches with the baseline file. 

```{r}
  anti_join(mci_data, amyloid_df, by = "id") 
```
There are 5 people that are in baseline list but not amyloid 

```{r}
  anti_join(amyloid_df, mci_data, by = "id")
```
There are 106 people that are in amyloid but not in baseline

```{r}
mci_combined = 
  inner_join(mci_data, amyloid_df, by = "id")
```

There are 381 observations and 11 variables left in the combined file.

```{r}
write_csv(mci_combined, "data_mci/mci_combined.csv")
```

